---
title: "pre_class_1_exercises"
author: "Kenneth C. Enevoldsen"
date: "7/7/2019"
output:   
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pre-class 1 Exercises

The warm-up for class will use the mtcars dataset which is always present in R.
The following code shows you how load the dataset

```{r load the data}
pacman::p_load(pacman, tidyverse) #laod packages
? mtcars # what are the variables
cars <- mtcars %>% rownames_to_column(var = "type") # tranform to tibble
```


This a brief re-introduction to the tidyverse package in R

This is something we went through during 1st semester so you should be able to do it, however if the summer have been to harsh I recommend consulting R for data science chapter 5: https://r4ds.had.co.nz
```{r}
# Using mutate - convert miles per gallon to liters
cars <- cars %>% mutate(mpl = mpg*3.78)

# Using filter - find all the cars which have more that 4 gears and are automatic
cars_4_auto <- cars %>% filter(am == 1|gear > 4)

# Using select - create a column dataframe with only miles pr. gallon, weight and number og gears
select <- select(cars, mpg, wt, gear)
?select

# Using arrange - What cars are most fuel efficient? (lowest miles pr. liter/gallon)
cars <- arrange(cars, mpg)

# Using Group_by and summarise - Find out how many cars have 3 gears, how many have 4 and how many have 5 (tip use n(), with summarise to count number of occurences)
sum <- cars %>% 
  group_by(gear) %>% 
  summarise(n())

# If you haven't yet, solve the above task using pipe, they look like this '%>%' and can be read as 'then' e.g. the following lines:
mtcars %>% 
  mutate(number = 200) %>% 
  summarise(sum_num = sum(number)) %>% 
  mutate(n_row = sum_num / 200)
# can be read as 'take mtcars, 
  # THEN add a column called numbers which is equal to 200
  # THEN the summarise the using the sum of numbers 
  # THEN divide the sum of the number by 200 to get the number of rows

#create some sort of plot plotting the data

<<<<<<< HEAD
ggplot(cars, aes(gear)) + geom_histogram(bins=10) + scale_x_continuous(breaks=c(3,4,5)) + xlab("number of gears") + ylab("amount of cars")
      
#does my pushing works??
=======
ggplot(cars, aes(gear)) +
  geom_histogram(bins=10) +
  scale_x_continuous(breaks=c(3,4,5)) +
  xlab("number of gears") +
  ylab("amount of cars")

ggplot(cars, aes(cyl, mpg)) +
  geom_point() +
  geom_smooth(method="lm")


>>>>>>> 802a6e3fa775ad6d4ae34bf64de043fbfebd0cdc
```


