---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.

Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?
- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

We first want to build a logistic regression to see whether you can diagnose schizophrenia from your best acoustic feature. Let's use the full dataset and calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve). You need to think carefully as to how we should (or not) use study and subject ID.

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures. Alternatively, the groupdata2 and cvms package created by Ludvig are an easy solution. 

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.N.N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?
N.N.N.N.B. A more advanced solution could rely on the tidymodels set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)

.N.N.N.B. A more advanced solution could rely on the tidymodels set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)


Data&packages, preprocessing data, models and their performance matrics
```{r}
### data&package ###

pacman::p_load(tidyverse, tidymodels, ggplot2, groupdata2, parsnip, randomForest, C50)
df <- read.csv(file="~/alexandersgitnest/Assignments/3/3.1/demo_art_pitch_danish")

data <- df
data

# deleting unnecessary columns (uID preserved)
data[,c(1,2,6,8,10,11,12,13,14,15,20,28,30,31,32,34)] <- NULL

#making outcome variabel factor
data$diagnosis<- as.factor(data$diagnosis)


### partioning into test and trainig ###

set.seed(5)#setting to have random random always the same

df_list <- partition(data, p = 0.2, cat_col = c("diagnosis"), id_col = c("uID"), list_out = T)
df_test = df_list[[1]]
df_train = df_list[[2]]

#saving id columns in the corresponding order to partitioned data for later analysis
df_test_uID <- df_test[,1]
df_test_uID

df_train_uID <- df_train[,1]
df_train_uID

#deleting id columns
df_test[,1] <- NULL
df_train[,1] <- NULL

# for (i in 1:4){
#   study=data$study %>% 
#     stats::filter(data,study==i)
#   opskrift <- study %>% recipe(diagnosis ~ .) %>% # defines the outcome
#   step_center(all_numeric()) %>%
#   step_scale(all_numeric()) %>% # scales numeric predictors
#   step_corr(all_numeric()) %>%
#   check_missing(everything()) %>%
#   prep(training = df_train)
#   return(opskrift)
# }
# opskrift (in danish recipe)
opskrift <- df_train %>% recipe(diagnosis ~ .) %>% # defines the outcome
  step_center(all_numeric()) %>%
  step_scale(all_numeric()) %>% # scales numeric predictors
  step_corr(all_numeric()) %>% 
  check_missing(everything()) %>%
  prep(training = df_train)


#inspecting what we have done
train_baked <- juice(opskrift) # extract df_train from rec
test_baked <- opskrift %>% bake(df_test)


### creating models using machine learning###

## linear regression
m_lr <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") %>%
  fit(diagnosis ~ . , data = train_baked)

## support vector machine
m_svm <-
  svm_rbf() %>%
  set_mode("classification") %>% 
  set_engine("kernlab") %>%
  fit(diagnosis ~ . , data = train_baked)

## random forest
m_rf <- 
  rand_forest() %>% 
  set_mode("classification") %>% 
  set_engine("randomForest") %>% 
  fit(diagnosis ~ . , data=train_baked)

## boost tree
m_bt <- 
  boost_tree() %>% 
  set_mode("classification") %>% 
  set_engine("C5.0") %>% 
  fit(diagnosis ~ . , data=train_baked)


### applying model to test set ###


## Predicting class
log_class <-  m_lr %>% 
  predict(new_data=test_baked) 
## prob. of class:
log_prop <- m_lr %>% 
  predict(new_data=test_baked, type = "prob") %>% 
  pull(.pred_1)


## getting probabilities of 4 models at once
test_results <- 
  test_baked %>% 
  select(diagnosis) %>% 
  mutate(
    lr_class = predict(m_lr, new_data = test_baked) %>% 
      pull(.pred_class),
    lr_prob  = predict(m_lr, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1),
    svm_class = predict(m_svm, new_data = test_baked) %>% 
      pull(.pred_class),
    svm_prob  = predict(m_svm, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1),
    rf_class = predict(m_rf, new_data = test_baked) %>% 
      pull(.pred_class),
    rf_prob  = predict(m_rf, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1),
    bt_class = predict(m_bt, new_data = test_baked) %>% 
      pull(.pred_class),
    bt_prob  = predict(m_bt, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1) # I left all pulls empty as it didn't work with any argument I tried
  )
## checking the probabilities
test_results


### Performance metrics ###

## getting accuracy, plotting cumulative gains curve, and roc curve

# linear regression
metrics(test_results, truth = diagnosis, estimate = lr_class) 

test_results %>% 
  mutate(lr_prob = 1 - lr_prob) %>% # for the plot to show correctly (otherwise the line would be flipped)
  gain_curve(truth = diagnosis, lr_prob) %>% 
  autoplot()

test_results %>%
  roc_curve(truth = diagnosis, lr_prob) %>% 
  autoplot()

# support vect machine
metrics(test_results, truth = diagnosis, estimate = svm_class) 

test_results %>% 
  mutate(svm_prob = 1 - svm_prob) %>% # for the plot to show correctly (otherwise the line would be flipped)
  gain_curve(truth = diagnosis, svm_prob) %>% 
  autoplot()

test_results %>%
  roc_curve(truth = diagnosis, svm_prob) %>% 
  autoplot()

# random forest
metrics(test_results, truth = diagnosis, estimate = rf_class) 

test_results %>% 
  mutate(rf_prob = 1 - rf_prob) %>% # for the plot to show correctly (otherwise the line would be flipped)
  gain_curve(truth = diagnosis, rf_prob) %>% 
  autoplot()

test_results %>%
  roc_curve(truth = diagnosis, rf_prob) %>% 
  autoplot()

# boost tree
metrics(test_results, truth = diagnosis, estimate = bt_class) 

test_results %>% 
  mutate(bt_prob = 1 - bt_prob) %>% # for the plot to show correctly (otherwise the line would be flipped)
  gain_curve(truth = diagnosis, bt_prob) %>% 
  autoplot()

test_results %>%
  roc_curve(truth = diagnosis, bt_prob) %>% 
  autoplot()



### diagnosing schizophrenia with the only best prediction - ... ###

## model
m_lr_best <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") %>%
  fit(diagnosis ~ speechrate..nsyll.dur. , data = train_baked)

##getting probabilities

test_best_results <- 
  test_baked %>% 
  select(diagnosis) %>% 
  mutate(
    lr_best_class = predict(m_lr_best, new_data = test_baked) %>% 
      pull(.pred_class),
    lr_best_prob  = predict(m_lr_best, new_data = test_baked, type = "prob") %>% 
      pull(.pred_1))

test_best_results

## performance metrics

metrics(test_best_results, truth = diagnosis, estimate = lr_best_class) 

test_best_results %>% 
  mutate(lr_best_prob = 1 - lr_best_prob) %>% # for the plot to show correctly (otherwise the line would be flipped)
  gain_curve(truth = diagnosis, lr_best_prob) %>% 
  autoplot()

test_best_results %>%
  roc_curve(truth = diagnosis, lr_best_prob) %>% 
  autoplot()

```



Multimodal cross-validation
```{r}
### Multimodal cross-validation ###

## appending the id column back to dataframes in order to create balanced folds
df_train <- mutate(df_train, uID=df_train_uID$uID)
df_test <- mutate(df_test, uID=df_test_uID$uID)

cv_folds <- vfold_cv(df_train, v = 10, repeats = 10, strata = diagnosis, group=uID)

cv_folds <- cv_folds %>% 
  mutate(recipes = splits %>%
           # prepper is a wrapper for `prep()` which handles `split` objects
           map(prepper, recipe = opskrift),
         train_data = splits %>% map(training))

# train model of each fold
  # create a non-fitted model
m_lr_nfit <- 
  logistic_reg() %>%
  set_mode("classification") %>% 
  set_engine("glm") 


cv_folds <- cv_folds %>%  mutate(
  log_fits = pmap(list(recipes, train_data), #input 
                            ~ fit(m_lr_nfit, formula(.x), data = bake(object = .x, new_data = .y)) # function to apply
                 ))

cv_folds %>% head(5)


predict_log <- function(split, rec, model) {
  # IN
    # split: a split data
    # rec: recipe to prepare the data
    # 
  # OUT
    # a tibble of the actual and predicted results
  baked_test <- bake(opskrift, testing(split))
  tibble(
    actual = baked_test$diagnosis,
    predicted = predict(model, new_data = baked_test) %>% pull(.pred_class),
    prop_diag =  predict(model, new_data = baked_test, type = "prob") %>% pull(.pred_1),
    prop_no_diag =  predict(model, new_data = baked_test, type = "prob") %>% pull(`.pred_0`)
  ) 
}

# apply our function to each split, which their respective recipes and models (in this case log fits) and save it to a new col
cv_folds <- cv_folds %>% 
  mutate(pred = pmap(list(splits, recipes, log_fits), predict_log))



## Evaluating performance metrics:
evaluation <- 
  cv_folds %>% 
  mutate(
    metrics = pmap(list(pred), ~ metrics(., truth = actual, estimate = predicted, prop_diag))) %>% 
  select(id, id2, metrics) %>% 
  unnest(metrics)

#inspect performance metrics
evaluation %>% 
  select(repeat_n = id, fold_n = id2, metric = .metric, estimate = .estimate) %>% 
  spread(metric, estimate) %>% 
  summary()

#mean 0.65, kap 0.30, mm_log_loss 0.96, roc_auc 0.68

```








